### 2019-06-24 thesis-template/msc-thesis.tex
update tests

\section{Light Transport Gradient (LTG)}
\label{sec:ltg}
Now, that we are able to propagate intensity in grid, directed by the eigenvectors of the tensor field, we aim to do something more elaborate to segment ROIs (regions of interest) characterised by divergence/convergence of tensor field lines and drastic changes in anisotropy. For that we imagine kind of a local and spatial gradient, as we place light sources in two close-by positions and eventually in two close-by directions and analyze the resulting final light distribution. We consider the light distributions and their difference image for the drain test field (cf. Fig. \ref{drain} in sect. \ref{chap:grundlagen}) to have an idea, for which kind or types of cases/configurations we obtain inherently different light distributions in particular. We choose to use this test case, because it constitutes an absolutely pure example par excellence for a FTLE computation exposing raw tensor field lines diverging.  Fig. \ref{evolution} depicts final light distributions (cf. sect. \ref{sec:ltg} Fig. \ref{rings-tests}) for placing the light source one step to the left and one to the right from the central axis. We can observe the light cones cancelling out each other in the center of the difference image. That is, per default, light distributions cancel each other out for the same light source position and direction (or for continous overlaps). In this case, for the \enquote{drain} test field, we have tensor field lines diverging leading to a strong difference for light sources in finite distance, which form these curves.

@@

The complementing idea for the angular domain would be placing the light source at the same position but pointing in differing directions:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_leftangle.png}
    \label{a)}
    \caption*{a) final light distribution for left angular offset}
  \end{minipage} $-$
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_rightangle.png}
    \label{b)}
    \caption*{b) final light distribution for right angular offset}
  \end{minipage} $=$
   \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_anglediff.png}
    \label{b)}
    \caption*{difference image for a) and b)}
  \end{minipage}
  \caption{``drain'' test field}
\label{evolution}
\end{figure}

@@

\begin{algorithm}
\caption{Propagate Light Distribution}\label{PLD}
\begin{algorithmic}[1]
\Procedure{propagateDist}{i,j,$\omega$}\Comment{Input position $j,i$ and direction $\omega$}
\State \Call{fill}{\textit{bufferA}, $0.0$} \Comment{Reset and Initialize}
\State \Call{fill}{\textit{bufferB}, $0.0$}
\State $\mathit{sumMem} \gets 0.0$
\State $\mathit{finished} \gets \mathit{false}$
\State $\mathit{index} \gets (j\cdot\mathit{width}+i)\cdot steps$ \Comment{compute $1D$ (cell) index}
\State $\mathit{bufferA(index,\omega)} \gets steps$ \Comment{write light src }
\While{$\Delta\Phi_{total} < \mathit{\epsilon }$ } \Comment{while convergence criterion not met..}
\State $\mathit{sumA} \gets 0.0$ \Comment{reset sum}
\State \Call{propagate}{\null} \Comment{propagate \textit{bufferA} (src) in \textit{bufferB} (tar)}
\State $\mathit{sumA} \gets$\Call{sum}{\textit{bufferB}}  \Comment{sum up energies}
\State $\Delta\Phi_{total} \gets \lvert\mathit{sumA}-\mathit{sumMem}\rvert $\Comment{compute difference to prev. iter.}
\State $\mathit{sumMem} \gets \mathit{sumA}$\Comment{save sum for next iter.}
\State \Call{swap}{\textit{bufferA}, \textit{bufferB}} \Comment{swap buffers for restart}
\State $\mathit{bufferA(index,\omega)} \gets steps$ \Comment{re-write light src}
\State \Call{fill}{\textit{bufferB}, $0.0$}
\If {$\mathit{ctr}>\mathit{limit}$}\Comment{stop on iteration limit}
\State break
\EndIf
\EndWhile\label{euclidendwhile} \Comment{final light distribution stored in bufferA..}
\State return \textit{bufferA}
\EndProcedure
\end{algorithmic}
\end{algorithm}
Note that we set the light source intensity to constant value $\mathit{steps}$ to require a mean intensity of $\mu_I=1.0$ comparable to the one of a unit circle independently of the angular resolution.

@@

\begin{algorithm}
\caption{Propagate BufferA to BufferB}\label{PB}
\begin{algorithmic}[1]
\Procedure{propagate}{}
 \For{each cell $c$ in bufferA}
\State $\mu_{I_c}\gets mean(I_c(\omega))$ \Comment{Compute Mean Intensity}
\If {$\mu_{I_c}=0.0$}\Comment{break on trivial null sample}
\State break
\EndIf
\State $\mu_{T_c}\gets mean(T_c(\omega))$ \Comment{Comp. Mean Transmission}
\State $\mu_{\mathit{T_cI_c}}\gets mean(T_c(\omega)\cdot I_c(\omega))$ \Comment{Comp. Mean Transmitted Intensity}
\State $n_c \gets \frac{\mu_{T_c}\mu_{I_c}}{\mu_{T_cI_c}}$\Comment{Comp. normalization factor for cur. cell}
 \For{each direction $k$ in $[0,7]$}\Comment{For all neighbors, do..}
 	\State $\gamma \gets k\frac{\pi}{4}$\Comment{Offset to compute cone center angle}
 	\If {$k\bmod 2 = 0$}\Comment{if even (face) neighbor..}
	\State $\mathit{energy} = \int_{\gamma-\pi/4}^{\gamma+\pi/4}n_c\epsilon_\alpha T_c(\omega)I_c(\omega)\mathop{d\omega}$
	\Else\Comment{if odd (diagonal) neighbor..}
	\State $\mathit{energy} = \int_{\gamma-\beta}^{\gamma+\beta}n_c\epsilon_\beta T_c(\omega)I_c(\omega)\mathop{d\omega}$
	\EndIf
	\State $\mathit{index} \gets c+\mathit{deltaIndex}(k)$\Comment{assign neighbor destination index}
	\State $\mathit{bufferB(index}\gets \mathit{energy}\cdot cos_k(\omega)+\mathit{bufferB(index}\cdot\mathit{steps)}$\\
	\Comment{Scale cosine lobe with summed energy and accumulate}\\
	\Comment{$y=a*x+y\rightarrow$ daxpy-OP}
 \EndFor
 \EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}

@@

\section{Light transport gradient (LTG)}
The light transport gradient approach, which is closely related to the FTLE approach in vector fields, is evaluated for some more synthetic and real datasets in this section. We first evaluate a classic textbook example the \enquote{Double Gyre}, which is a quadruple gyre in our case for formatting reasons, yet yielding the same conceptual results. We compare the glyph representation to the LTG magnitude for a single slice ($0^\circ$) again:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[width=\textwidth]{img/gyre.png}
    \caption{\enquote{Gyre} test field}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_ftle.PNG}
     \caption{LTG on \enquote{Gyre}($0^\circ$)}
    \label{b)}
  \end{minipage}
\label{drain_contour}
\end{figure}
We can observe the method segmenting ridges in local ROIs exhibiting diverging tensor field lines w.r.t. the $x$-axis ($0^\circ$). Note that even regions in antiparallel directions ridgres are detected to (literally) some degree: Remark the ridges being contiguos lines at the level of the gyre centers where converging turns into diverging. We again depict volume representations for this example:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[width=\textwidth]{img/gyre_volume.png}
    \caption{IsoVolume on \enquote{Gyre} test field (low thresh)}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_ftle-highvolume.PNG}
     \caption{IsoVolume on \enquote{Gyre}(high thresh)}
    \label{b)}
  \end{minipage}
\label{gyre_volume}
\end{figure}
When we consider the volumes for the \enquote{gyre} test field, we can remark the holes, which represent the gyres in the centers of the volume for a low and a high threshold, whereby we extract (filter out) the previously observed ridges in $3D$ for all directions $\omega$ simultaneously.

Now, that it passed the textbook examples, we push it forward to be applied in real (measured) data to show its performance in discerning and segmenting ROIs from a human brain. The approach will detect LCS in the brain's tissue from MRI-measured diffusion tensors of a male subject at the age of 27. We compare t

>>>>

\section{Light Transport Gradient (LTG)}
\label{sec:ltg}
Now, that we are able to propagate intensity in grid, directed by the eigenvectors of the tensor field, we aim to do something more elaborate to segment ROIs (regions of interest) characterised by divergence/convergence of tensor field lines and drastic changes in anisotropy. For this purpose, we imagine kind of a local and spatial gradient, as we place light sources in two close-by positions and eventually in two close-by directions and analyze the resulting final light distribution. We consider the light distributions and their difference image for the drain test field (cf. Fig. \ref{drain} in sect. \ref{chap:grundlagen}) to have an idea, for which kind or types of cases/configurations we obtain inherently different light distributions in particular. We choose to use this test case, because it constitutes an absolutely pure example par excellence for a FTLE computation exposing raw tensor field lines diverging.  Fig. \ref{evolution} depicts final light distributions (cf. sect. \ref{sec:ltg} Fig. \ref{rings-tests}) for placing the light source one step to the left and one to the right from the central axis. We can observe the light cones cancelling out each other in the center of the difference image. That is, per default, light distributions cancel each other out for two close-by light source positions or directions (or more specifically: for continous overlaps in their distributions). In this case, for the \enquote{drain} test field, we have tensor field lines diverging leading to a strong difference for light sources in finite distance, which ín turn form these curves depicted in Fig. \ref{evolution}.

@@

The complementing idea for the angular domain would be: placing the light source at the same position but pointing in differing directions.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_leftangle.png}
    \label{a)}
    \caption*{a) final light distribution for left angular offset}
  \end{minipage} $-$
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_rightangle.png}
    \label{b)}
    \caption*{b) final light distribution for right angular offset}
  \end{minipage} $=$
   \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth]{img/ftle_anglediff.png}
    \label{b)}
    \caption*{difference image for a) and b)}
  \end{minipage}
  \caption{``drain'' test field}
\label{evolution}
\end{figure}

@@

\begin{algorithm}
\caption{Propagate Light Distribution}\label{PLD}
\begin{algorithmic}[1]
\Procedure{propagateDist}{i,j,$\omega$}\Comment{Input position $j,i$ and direction $\omega$}
\State \Call{fill}{\textit{bufferA}, $0.0$} \Comment{Reset and Initialize}
\State \Call{fill}{\textit{bufferB}, $0.0$}
\State $\mathit{sumMem} \gets 0.0$
\State $\mathit{finished} \gets \mathit{false}$
\State $\mathit{index} \gets (j\cdot\mathit{width}+i)$ \Comment{compute $1D$ (cell) index}
\State $\mathit{bufferA(index,\omega)} \gets steps$ \Comment{write light src }
\While{$\Delta\Phi_{total} < \mathit{\epsilon }$ } \Comment{while convergence criterion not met..}
\State $\mathit{sumA} \gets 0.0$ \Comment{reset sum}
\State \Call{propagate}{\null} \Comment{propagate \textit{bufferA} (src) in \textit{bufferB} (tar)}
\State $\mathit{sumA} \gets$\Call{sum}{\textit{bufferB}}  \Comment{sum up energies}
\State $\Delta\Phi_{total} \gets \lvert\mathit{sumA}-\mathit{sumMem}\rvert $\Comment{compute difference to prev. iter.}
\State $\mathit{sumMem} \gets \mathit{sumA}$\Comment{save sum for next iter.}
\State \Call{swap}{\textit{bufferA}, \textit{bufferB}} \Comment{swap buffers for restart}
\State $\mathit{bufferA(index,\omega)} \gets steps$ \Comment{re-write light src}
\State \Call{fill}{\textit{bufferB}, $0.0$}
\If {$\mathit{ctr}>\mathit{limit}$}\Comment{stop on iteration limit}
\State break
\EndIf
\EndWhile\label{euclidendwhile} \Comment{final light distribution stored in bufferA..}
\State return \textit{bufferA}
\EndProcedure
\end{algorithmic}
\end{algorithm}
Note that we set the light source intensity to constant value $\mathit{steps}$ to require a mean intensity of $\mu_I=1.0$ comparable to the one of a unit circle independently of the angular resolution.

@@

\begin{algorithm}
\caption{Propagate BufferA to BufferB}\label{PB}
\begin{algorithmic}[1]
\Procedure{propagate}{}
 \For{each cell $c$ in bufferA}
\State $\mu_{I_c}\gets mean(I_c(\omega))$ \Comment{Compute Mean Intensity}
\If {$\mu_{I_c}=0.0$}\Comment{break on trivial null sample}
\State break
\EndIf
\State $\mu_{T_c}\gets mean(T_c(\omega))$ \Comment{Comp. Mean Transmission}
\State $\mu_{\mathit{T_cI_c}}\gets mean(T_c(\omega)\cdot I_c(\omega))$ \Comment{Comp. Mean Transmitted Intensity}
\State $n_c \gets \frac{\mu_{T_c}\mu_{I_c}}{\mu_{T_cI_c}}$\Comment{Comp. normalization factor for cur. cell}
 \For{each direction $k$ in $[0,7]$}\Comment{For all neighbors, do..}
 	\State $\gamma \gets k\frac{\pi}{4}$\Comment{Offset to compute cone center angle}
 	\If {$k\bmod 2 = 0$}\Comment{if even (face) neighbor..}
	\State $\mathit{energy} = \int_{\gamma-\pi/4}^{\gamma+\pi/4}n_c\epsilon_\alpha T_c(\omega)I_c(\omega)\mathop{d\omega}$
	\Else\Comment{if odd (diagonal) neighbor..}
	\State $\mathit{energy} = \int_{\gamma-\beta}^{\gamma+\beta}n_c\epsilon_\beta T_c(\omega)I_c(\omega)\mathop{d\omega}$
	\EndIf
	\State $\mathit{index} \gets c+\mathit{deltaIndex}(k)$\Comment{assign neighbor destination index}
	\State $\mathit{bufferB(index)}\gets \mathit{energy}\cdot cos_k(\omega)+\mathit{bufferB(index}$\\
	\Comment{Scale cosine lobe with summed energy and accumulate}\\
	\Comment{$y=a*x+y\rightarrow$ daxpy-OP}
 \EndFor
 \EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}

@@

\section{Light Transport Gradient (LTG)}
The light transport gradient approach, which is closely related to the FTLE approach in vector fields, is evaluated for some more synthetic and real datasets in this section. We first evaluate a classic textbook example called the \enquote{Double Gyre} \cite{shawn}, which is a quadruple gyre in our case for formatting reasons, yet yielding the same conceptual results. We compare the glyph representation to the LTG magnitude for a single slice ($0^\circ$) again:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[width=\textwidth]{img/gyre.png}
    \caption{\enquote{Gyre} test field}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_ftle.PNG}
     \caption{LTG on \enquote{Gyre}($0^\circ$)}
    \label{b)}
  \end{minipage}
\label{drain_contour}
\end{figure}
We can observe the method segmenting ridges in local ROIs exhibiting diverging tensor field lines w.r.t. the $x$-axis ($0^\circ$). Note that even regions in antiparallel directions ridgres are detected to (literally) some degree: Remark the ridges being contiguos lines at the level of the gyre centers where converging turns into diverging. Also remember, that the resulting ridges are not to be concerned sharp ones, since we measure the separation of whole light distributions, which should be considered a continous response or measure. We again depict volume representations for this example:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[width=\textwidth]{img/gyre_volume.png}
    \caption{IsoVolume on \enquote{Gyre} test field (low thresh)}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[height=\textwidth]{img/gyre_ftle-highvolume.PNG}
     \caption{IsoVolume on \enquote{Gyre}(high thresh)}
    \label{b)}
  \end{minipage}
\label{gyre_volume}
\end{figure}
When we consider the volumes for the \enquote{gyre} test field, we can remark the holes, which represent the gyres in the centers of the volume for a low and a high threshold, whereby we extract (filter out) the previously observed ridges in $3D$ for all directions $\omega$ simultaneously.

Now, that it passed the textbook examples, we push it forward to be applied in real (measured) data to show its performance in discerning and segmenting ROIs from a human brain. The approach will detect LCS in the brain's tissue from MRI-measured diffusion tensors of a male subject at the age of 27. We compare ParaView's glyph representation to the FTLE method LTG for the test field \enquote{brain}:
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[width=\textwidth]{img/brain_ellipsoid_glyphs.PNG}
    \caption{\enquote{Brain} test field glyph representation}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[height=\textwidth]{img/brain_ftle.PNG}
     \caption{LTG on \enquote{Brain} test field}
    \label{b)}
  \end{minipage}
\label{gyre_volume}
\end{figure}
Regarding Fig. \ref{gyre_volume}, we find that the approach detects the boundaries of concavities (lat.: Gyri) in the brain, forming the lobus cerebri (commonly known a lobes). The invented method could thus ,e.g. , aid to compute $3D$ models of a subject in medical sciences. It emphasizes regions with much flow-rate (throughput), because in this areas the anisotropy and/or tensor magnitude is increased, preventing the intensities from being absorbed, hence yielding to a high FTLE-response.
